What's in a Name? \\ Linear Temporal Logic \\ Literally Represents Time Lines

### 2023-08-27 paper/main.tex
typo fixes

\begin{definition}[$\omega$-regular expression]\label{def:omega-regex}
    Regular expression concerns only finite-length strings. However, since \ltl\ formulas reason about events that happen over an infinite-length timeline, we need to model it using infinite regular expressions (i.e. $\omega$-regular expression). $\omega$-regular expression is defined by the following grammar:
    \begin{align*}
        B ::= A^{\omega} \mid AB \mid B + B
    \end{align*}
\end{definition}
\begin{definition}[Semantics of $\omega$-regular expression]\label{def:omega-semantics}
    Let $\Sigma^{\omega}$ denote the set of infinite-length strings over fixed alphabet $\Sigma$. Let $L_{\omega}(B)$ denote the $\omega$-language accepted by $\omega$-regular expression $B$. Then $L_{\omega}(B)$ is inductively defined as
    \begin{align*}
        L_{\omega}(A^{\omega}) & = \{s_1s_2s_3\cdots \mid s_i \in L(A) \text{ and } i \ge 1\} \tag{$\epsilon
         \not\in L(A)$} \\
        L_{\omega}(AB) & = \{s_1s_2 \mid s_1 \in L(A) \text{ and } s_2 \in L_{\omega}(B)\} \\
        L_{\omega}(B_1 + B_2) &= L_{\omega}(B_1) \cup L_{\omega}(B_2)
    \end{align*}
\end{definition}

@@

\subsection{LTL to BA} \label{ltl2aut}
The automata-theoretic approach~\cite{ORBi-a8d60ab6-1101-4434-9511-c01ea4e5a15b} of studying \ltl\ (i.e. reducing \ltl\ formulas to their corresponding \Buchi automata) has been well-studied. Many algorithms and optimizations have been studied such as~\cite{DGV99, F03}. Our tool \tool uses SPOT~\cite{Dur22}\footnote{For our tool, we use SPOT version 2.11.} for this step, which provides an implementation of the translation from \ltl\ formulas to \Buchi automata.

@@

\begin{algorithm}[h!]
    \caption{nfa2regex}
    \begin{algorithmic}
        \Require ($G, s, f$), an NFA with initial state $s$ and final state $f$
        \Ensure The regular expression corresponding to all paths from $s$ to $f$
        % WHile there exists any edge except r_sf
        \While{ there exists an interior vertex $v$ }
            \State reduce\_nfa($G, v$)
            \State combine multi-edges, i.e. $r1 : u \to w, r2 : u \to w$  to $r1 | r2 : u \to w$
        \EndWhile
        \State \Return $(r_\mathrm{ss}| r_\mathrm{sf} r_\mathrm{ff}^* r_\mathrm{fs})^* r_\mathrm{sf} r_\mathrm{ff}$
        % if called by first visit, this is r_\mathrm{ss}* r_\mathrm{sf}
        % if called by ba2wregex, s=f, this is r_\mathrm{ff}
        % consider modularizing fns differently
        % \State \Return regexp labelling the edge $s\to f$
    \end{algorithmic}
\end{algorithm}

@@

\section{Discussion}
Based on our results, we propose the following questions for discussions and future works:
\begin{itemize}
    \item Can there be more simplifications methods as described in Section~\ref{regex-simplify}? The answer is definitely positive. On one hand, more rule-based simplifications can be discovered by finding more patterns in the generated $\omega$-regular expressions. On the other hand, other simplification are possible: for example, if the generated $\omega$-regular expression is $r_1 + r_2$ where $r_1, r_2$ are not syntactically equivalent (or equivalent up to algebraic rules) $\omega$-regular expressions, but they accepts the same set of infinite-length strings, then we can simplify this down to just $r_1$. This reduce the problem to equivalence test on $\omega$-regular expressions, which is decidable. In the similar vein, many more simplifications can be used, but the question is how efficient are they? In the end, one have to balance trade-off between the efficiency of the tool and the simplicity of the result.
    %\item Our tool is effective at translating a wide range of LTL formulas. Even with LTL formulae with complex nesting structure are represented by relatively readable timeline visualizations.
    \item Can we define timeline more formally so that one can do this process reversely (i.e. a tool that converts timeline representation to its corresponding \ltl\ formula)? That would be a very useful tool in practice. However, a direct reverse of the algorithm we presented in this paper may not be viable.
\end{itemize}

>>>>

\begin{definition}[$\omega$-regular expression]\label{def:omega-regex}
    Regular expression concerns only finite-length strings. However, since \ltl\ formulas reason about events that happen over an infinite-length timeline, we need to model it using infinite regular expressions (i.e., $\omega$-regular expression). $\omega$-regular expression is defined by the following grammar:
    \begin{align*}
        B ::= A^{\omega} \mid AB \mid B + B
    \end{align*}
\end{definition}
\begin{definition}[Semantics of $\omega$-regular expression]\label{def:omega-semantics}
    Let $\Sigma^{\omega}$ denote the set of infinite-length strings over fixed alphabet $\Sigma$. Let $L_{\omega}(B)$ denote the $\omega$-language accepted by $\omega$-regular expression $B$. Then $L_{\omega}(B)$ is inductively defined as
    \begin{align*}
        L_{\omega}(A^{\omega}) & = \{s_1s_2s_3\cdots \mid s_i \in L(A) \text{ and } i \ge 1\} \tag{$\epsilon
         \not\in L(A)$} \\
        L_{\omega}(AB) & = \{s_1s_2 \mid s_1 \in L(A) \text{ and } s_2 \in L_{\omega}(B)\} \\
        L_{\omega}(B_1 + B_2) &= L_{\omega}(B_1) \cup L_{\omega}(B_2)
    \end{align*}
\end{definition}

@@

\subsection{LTL to BA} \label{ltl2aut}
The automata-theoretic approach~\cite{ORBi-a8d60ab6-1101-4434-9511-c01ea4e5a15b} of studying \ltl\ (i.e., reducing \ltl\ formulas to their corresponding \Buchi automata) has been well-studied. Many algorithms and optimizations have been studied such as~\cite{DGV99, F03}. Our tool \tool uses SPOT~\cite{Dur22}\footnote{For our tool, we use SPOT version 2.11.} for this step, which provides an implementation of the translation from \ltl\ formulas to \Buchi automata.

@@

\begin{algorithm}[h!]
    \caption{nfa2regex}
    \begin{algorithmic}
        \Require ($G, s, f$), an NFA with initial state $s$ and final state $f$
        \Ensure The regular expression corresponding to all paths from $s$ to $f$
        % WHile there exists any edge except r_sf
        \While{ there exists an interior vertex $v$ }
            \State reduce\_nfa($G, v$)
            \State combine multi-edges, i.e., $r1 : u \to w, r2 : u \to w$  to $r1 | r2 : u \to w$
        \EndWhile
        \State \Return $(r_\mathrm{ss}| r_\mathrm{sf} r_\mathrm{ff}^* r_\mathrm{fs})^* r_\mathrm{sf} r_\mathrm{ff}$
        % if called by first visit, this is r_\mathrm{ss}* r_\mathrm{sf}
        % if called by ba2wregex, s=f, this is r_\mathrm{ff}
        % consider modularizing fns differently
        % \State \Return regexp labelling the edge $s\to f$
    \end{algorithmic}
\end{algorithm}

@@

\section{Discussion}
Based on our results, we propose the following questions for discussions and future work:
\begin{itemize}
    \item Can there be more simplifications methods as described in Section~\ref{regex-simplify}? The answer is definitely positive. On one hand, more rule-based simplifications can be discovered by finding more patterns in the generated $\omega$-regular expressions. On the other hand, other simplification are possible: for example, if the generated $\omega$-regular expression is $r_1 + r_2$ where $r_1, r_2$ are not syntactically equivalent (or equivalent up to algebraic rules) $\omega$-regular expressions, but they accept the same set of infinite-length strings, then we can simplify this down to just $r_1$. This reduces the problem to equivalence test on $\omega$-regular expressions, which is decidable. In a similar vein, many more simplifications may apply, but the question is how \empu{useful} are they? In the end, we have to balance the efficiency, simplicity, and understandabillity of the visualization.
    %\item Our tool is effective at translating a wide range of LTL formulas. Even with LTL formulae with complex nesting structure are represented by relatively readable timeline visualizations.
    \item Can we define timeline more formally so that one can do this process reversely (i.e., a tool that converts timeline representation to its corresponding \ltl\ formula)? That would be a very useful tool in practice. However, a direct reverse of the algorithm we presented in this paper may not be viable.
\end{itemize}
